Flatiron analysis

### 2021-03-17 Rmd/Report analysis 03172021.Rmd
New Rmd Flatiron analysis

---
title: "Flatiron data cleaning and analysis report"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: flatly
    highlight: pygments
    df_print: paged
---

@@

# Data cleaning
***
<br>

@@

## Creatinine cleaning

Only creatinine from blood and serum are being used. All units are correct (mg/dL).  
Here an sample of the non-cleaned data with really low `testresultcleaned`.

```{r creatinine, rows.print=15, max.print=15}
creatinine %>% filter(str_detect(labcomponent, "blood|serum")) %>% select(-c(loinc, test, labcomponent)) %>% 
arrange(testresultcleaned)
```
All `testresultcleaned` and `testresult` variable are cleaned as :  
- creat < 0.8 was set at 0.8  
- creat > 3 are eliminated  
To create the final `creatinine` variable ,`testresultcleaned` and `testresult` are both used but `testresultcleaned` was used over `testresult` when both are available.

Here a sample of the cleaned data
```{r creatinine1, rows.print=15, max.print=15}
creatinine <- creatinine %>%
  # remove creatine urine
  filter(str_detect(labcomponent, "blood|serum")) %>% 
  mutate(creatinine = case_when(
    testresultcleaned < 0.8        ~ 0.8,
    testresultcleaned > 3          ~ NA_real_,
    TRUE                           ~ testresultcleaned
  )) %>% 
  # F092F8B827DDC kept umol/L as they actually are mg/dL range 45 to 90 μmol/L (0.5 to 1.0 mg/dL) for women.
  mutate(creatinine1 = case_when(
    testresult < 0.8               ~ 0.8,
    testresult > 3                 ~ NA_real_,
    TRUE                           ~ testresult
  )) %>% 
  mutate(creatinine = coalesce(creatinine, creatinine1)) %>% 
  filter(!is.na(creatinine)) %>% 
  mutate(testdate = as.Date(testdate, format = "%m/%d/%y")) %>% 
  select(c("patientid", creat_date = "testdate", "creatinine", creat_units = "testunitscleaned"))
creatinine %>% distinct(patientid, .keep_all = TRUE)

vitals <- vitals %>% 
  mutate(testdate = as.Date(testdate, format = "%m/%d/%y"))
body_surface_area <- vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)")
```
`r emo::ji("check")`
<br>

## Body surface area cleaning

`testresultcleaned` is all within expected range but only few data is available (`r NROW(body_surface_area %>% filter(!is.na(testresultcleaned)) %>% distinct(patientid, .keep.all=TRUE))` patients). Probably even less with an associated date.

```{r bsa}
body_surface_area %>% select(-c(loinc, test, labcomponent))
```

By using `testresult` and before cleaning only `r NROW(body_surface_area %>% filter(!is.na(body_surface_area$testresult)) %>% distinct(patientid, .keep.all=TRUE))` patients have data for their BSA. This was improved by using the Du Bois formula to calculate more BSA value as you will see later.

To create the final `BSA` variable ,`testresultcleaned` was used over `testresult` when both are available.  
`testresult` was further cleaned as data < 0.418 and > 4.14 are eliminated. These values were calculated from the biggest and the smallest individuals recorded.  
I also remove the outliers within patient (2 examples here where we have outliers).

```{r}
vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)" & patientid == "F2D5A8748B268") %>% 
  select(patientid, testdate, testresult, testresultcleaned)
vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)" & patientid == "F875EA1B044F0") %>% 
  select(patientid, testdate, testresult, testresultcleaned)
# F3AE4546B5D12
```

Here the cleaned data

```{r bsa1}
body_surface_area <- vitals %>% 
  filter(labcomponent == "Body Surface Area (BSA)") %>% # only 399 patients data
  # Limit max = 4.14, min = 0.418 calculated for biggest to smallest person
  mutate(testresult = ifelse(testresult < 0.418 | testresult > 4.14, NA_real_, testresult)) %>% 
  group_by(patientid) %>%
  # I used the median to remove outlier within a patient
  mutate(median_testresult = median(testresult)) %>% 
  ungroup() %>% 
  mutate(BSA = case_when(
    testresult > (median_testresult + 0.4) |
      testresult < (median_testresult - 0.4)      ~ NA_real_,
    TRUE                                          ~ testresult
  )) %>% 
  mutate(BSA = coalesce(testresultcleaned, BSA)) %>% 
  filter(!is.na(BSA)) %>% 
  mutate(BSA_units = "m2") %>% 
  select(c("patientid", bsa_date = "testdate", "BSA", "BSA_units"))

body_surface_area %>% distinct(patientid, .keep_all = TRUE) %>% arrange(patientid, bsa_date)
```
`r emo::ji("check")`
<br>

## Height

Height data is pretty complicated. Again I used `testresultcleaned` and `testresult` with `testresultcleaned` which was prioritized.  
`testresultcleaned` has units in cm but some data are wrong. `testresult` has results in cm, ft or in. It was pretty easy to clean data between these 3 units because the measurement cannot overlap.
Both are cleaned as < 120 or > 231 are eliminated. These value are for the shortest and tallest women recorded.
The units can be filled by patients when absent.

```{r height}
vitals %>% 
  filter(test == "body height" & patientid == "FCB6DAAFF6527") %>% select(-c(loinc, test, labcomponent, resultdate))
# FC109A6F719E7
# FC106805F9AB9
# F91590CB83EC8 weird
```

Outliers are eliminated as much as possible.

```{r height1}
vitals %>% 
  filter(test == "body height") %>% 
  # To get more data, I have data from testresult that can be used
  # But need to fill up testunits, transfer in kg and coalesce.
  group_by(patientid) %>% 
  fill(testunits, .direction = "downup") %>% # F004D683A1695 have testresult, F615C37BF1470 is wrong, Otherwise I checked them all to see we can do that
  ungroup() %>% 
  # Clean up outliers compare to tallest ans smallest record
  mutate(testresultcleaned = case_when(
    testresultcleaned > 231 | # Tallest women height
      testresultcleaned < 120 # Average height dwarf = 122 cm, 4 feet, I take 120 by loooking at the data
                                     ~ NA_real_,
    TRUE                             ~ testresultcleaned
  )) %>% 
  mutate(testresult = case_when(
    testresult > 231 & testunits == "cm" |
      testresult < 100 & testunits == "cm" |
      testresult > 7.58 & testunits == "ft" |
      testresult < 3.28 & testunits == "ft" |
      testresult > 90.94 & testunits == "in" |
      testresult < 39.37 & testunits == "in"|
      testresult == 0
                                     ~ NA_real_,
    TRUE                             ~ testresult
  )) %>% 
  # Clean up outliers within patient by using the median for each patients/unit. Eliminate the ones with too much variation.
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned)) %>% 
  ungroup() %>% 
  mutate(testresultcleaned_verified = case_when(
    testresultcleaned > (median_testresultcleaned + 5) | # choose 5 cm by default
      testresultcleaned < (median_testresultcleaned - 5)      ~ NA_real_, # F01413C06D921 weird
    TRUE                                                      ~ testresultcleaned
  )) %>% 
  filter(patientid == "F01413C06D921") %>% 
  select(c("patientid", testdate, testresultcleaned, testresultcleaned_verified,median_testresultcleaned, 
           testunitscleaned, testunits, testresult))
```
`r  emo::ji("question_mark")`
PB: Some patients have height but don't have dates associated. If all dates are `NA` for a patient, I cannot combined a specific `height` with a closest date to `episodedate` of the drug data.
Should I take a mean of all the height values measured or the first one - If I have to choose I would take the mean (patients shouldn't grow up or shorten much)?

Here is the data.
```{r height2}
height <- vitals %>% 
  filter(test == "body height") %>% 
  # To get more data, I have data from testresult that can be used
  # But need to fill up testunits, transfer in kg and coalesce.
  group_by(patientid) %>% 
  fill(testunits, .direction = "downup") %>% # F004D683A1695 have testresult, F615C37BF1470 is wrong, Otherwise I checked them all to see we can do that
  ungroup() %>% 
  # Clean up outliers compare to tallest ans smallest record
  mutate(testresultcleaned = case_when(
    testresultcleaned > 231 | # Tallest women height
      testresultcleaned < 120 # Average height dwarf = 122 cm, 4 feet, I take 120 by loooking at the data
                                     ~ NA_real_,
    TRUE                             ~ testresultcleaned
  )) %>% 
  mutate(testresult = case_when(
    testresult > 231 & testunits == "cm" |
      testresult < 100 & testunits == "cm" |
      testresult > 7.58 & testunits == "ft" |
      testresult < 3.28 & testunits == "ft" |
      testresult > 90.94 & testunits == "in" |
      testresult < 39.37 & testunits == "in"|
      testresult == 0
                                     ~ NA_real_,
    TRUE                             ~ testresult
  )) %>% 
  # Clean up outliers within patient by using the median for each patients/unit. Eliminate the ones with too much variation.
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned)) %>% 
  ungroup() %>% 
  mutate(testresultcleaned_verified = case_when(
    testresultcleaned > (median_testresultcleaned + 5) | # choose 5 cm by default
      testresultcleaned < (median_testresultcleaned - 5)      ~ NA_real_, # F01413C06D921 weird
    TRUE                                                      ~ testresultcleaned
  )) %>% 
  # select(c("patientid", testdate, testresultcleaned, median_testresultcleaned, testresultcleaned_verified, 
  #          testunitscleaned, testunits, testresult)) %>%

  group_by(patientid, testunits) %>%
  mutate(median_testresult = median(testresult)) %>% 
  ungroup() %>% 
  mutate(testresult_verified = case_when(
    testresult > (median_testresult + 3) & 
      testunits == "in" |
      testresult < (median_testresult - 3) & 
      testunits == "in"                         ~ NA_real_,
    testresult > (median_testresult + 5) & 
      testunits == "cm" |
      testresult < (median_testresult - 5) & 
      testunits == "cm"                         ~ NA_real_,
    TRUE                                        ~ testresult
  )) %>% 
  # select(c("patientid", testdate, testresult, median_testresult, testresult_verified,
  #          testunits, testunitscleaned, testresultcleaned, testresultcleaned_verified))

  mutate(height = case_when(
    !is.na(testresultcleaned_verified)          ~ round(testresultcleaned_verified / 100, 3), # cm to m
    testunits == "cm"                           ~ round(testresult_verified / 100, 3),
    testunits == "ft"                           ~ round(testresult_verified / 3.281, 3), # ft to m
    testunits == "in" | 
      !is.na(testresult)                        ~ round(testresult_verified / 39.37, 3) # in to m
    )) %>% 
  filter(!is.na(height)) %>% 
  mutate(height_units = "m") %>% 
  select(c("patientid", height_date = "testdate", "height", "height_units"))

height %>% distinct(patientid, .keep_all = TRUE)
```
Beside my question it is `r emo::ji("check")`

<br>

## Weight

The more complex cleaning (before the drugs...`r  emo::ji("loudly_crying_face")`). Contrary to height with units which don't overlap, weight has data in `kg` in `testresultcleaned` and `kg`, `lb`, `oz` in `testresult` which are overlapping (a person can be 150lbs or 150kg).  
Cleanup of `testresultcleaned`...  
```{r}
vitals %>%
  filter(test == "body weight") %>% 
  select(c("patientid", "testdate", "testunitscleaned", "testresultcleaned")) %>%
  # 1.Clean testresultcleaned
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(weight1 = case_when(
    testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
      testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
      (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
         testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
    TRUE ~ NA_real_
  )) %>% filter(patientid == "F0027D3926C88")
```
... and eliminate the outliers (a patient gaining 36.4 kg can be normal for FF5268589DBE2, 39.5kg F54F47F4419E0) but 37.7 doesn't look normal F4A3B1B659865, F6CA354D599AA. My present cleaning save some data but not everything...  
I just had an idea on how to improve that better while doing the report.  
`r emo::ji("cross_mark")``r emo::ji("light_bulb")``r emo::ji("bookmark")`
```{r}
vitals %>%
  filter(test == "body weight") %>% 
  select(c("patientid", "testdate", "testunitscleaned", "testresultcleaned")) %>%
  # 1.Clean testresultcleaned
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(weight1 = case_when(
    testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
      testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
      (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
         testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
    TRUE ~ NA_real_
  )) %>% filter(str_detect(patientid, "F6CA354D599AA|F4A3B1B659865|FF5268589DBE2|F93CEAE4BF764|F1493F8F5A924|F18E68D19D10F|F54F47F4419E0"))
```

I had to clean `testresult` for each units separately within a patients because I saw patients with 2 units in their records.
I also rescued the units when some are `NA` but not all and eliminate the outliers.

```{r}
# vitals %>%
#   filter(test == "body weight") %>% 
#   select(c("patientid", "testdate", "testunits", "testresult", "testunitscleaned", "testresultcleaned")) %>%
#   # 1.Clean testresultcleaned
#   group_by(patientid) %>%
#   mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
#   ungroup() %>%
#   mutate(weight1 = case_when(
#     testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
#       testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
#       (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
#          testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
#     TRUE ~ NA_real_
#   )) %>% 
#   # mutate(difference = (median_testresultcleaned)-testresultcleaned) # Can improve by > 30 and > 10 from previous measure so need to create a gap var ---------
#   # 2.Clean testresult to be used to find max and min
#   mutate(testresult = case_when(
#     testunits == "lb" &
#       testresult > 12 &
#       testresult < 643            ~ testresult,
#     testunits == "kg" &
#       testresult >= 5.5 &
#       testresult < 292            ~ testresult,
#     testunits == "oz" &
#       testresult > 194 &
#       testresult < 10286          ~ testresult,
#     is.na(testunits) & 
#     testresult > 5.5 & 
#       testresult < 10286          ~ testresult, # combine oz, kg, lbs for when units is NA
#     TRUE                          ~ NA_real_
#   )) %>% filter(str_detect(patientid, "F0027D3926C88")) %>% 
#   select(c("patientid", "testdate", "testunits", "testresult",)

weight <- vitals %>%
  filter(test == "body weight") %>% 
  select(c("patientid", "testdate", "testunits", "testresult", "testunitscleaned", "testresultcleaned")) %>%
  # 1.Clean testresultcleaned
  group_by(patientid) %>%
  mutate(median_testresultcleaned = median(testresultcleaned, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(weight1 = case_when(
    testresultcleaned >= 27 &       # Biggest recorded 291.6 kg (642.9 lbs, 10286 oz)
      testresultcleaned < 292 &     # lightest 5.5 kg (12 lbs, 194 oz)  F0027D3926C88
      (testresultcleaned > (median_testresultcleaned - 39.6) & # F1493F8F5A924, F54F47F4419E0, F18E68D19D10F
         testresultcleaned < (median_testresultcleaned + 39.6))     ~ testresultcleaned, 
    TRUE ~ NA_real_
  )) %>% 
  # mutate(difference = (median_testresultcleaned)-testresultcleaned) # Can improve by > 30 and > 10 from previous measure so need to create a gap var ---------
  # 2.Clean testresult to be used to find max and min
  mutate(testresult = case_when(
    testunits == "lb" &
      testresult > 12 &
      testresult < 643            ~ testresult,
    testunits == "kg" &
      testresult >= 5.5 &
      testresult < 292            ~ testresult,
    testunits == "oz" &
      testresult > 194 &
      testresult < 10286          ~ testresult,
    is.na(testunits) & 
    testresult > 5.5 & 
      testresult < 10286          ~ testresult, # combine oz, kg, lbs for when units is NA
    TRUE                          ~ NA_real_
  ))
```
```{r}
weight %>% 
  # 3. Calculate median for each units and remove data outside range for each patients/unit
  # 3. help distinguishing lbs to kg for same patient when unit is NA
  group_by(patientid, testunits) %>%
  mutate(median_testresult = median(testresult, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(testresult_verified = case_when(       # Can improve by > 30 and > 10 from previous measure so need to create a gap var --------
    testunits == "kg" &
    (testresult > (median_testresult + 38.5) | # choose 38.5 cm by default could do 32 if not for F534F9C4FF968
       # 12 F2F31607FD6AD , FA7C8143AEE58 gain 37kg, F847A533FAE7E gain 38, FBE6E66AB2C95 gain 41, FC09E942CB0D7 gain 49
       # F5B18C17A28F5 and F98077E9119EC FA53D0F672C7D FCFF0EDFE278F, FEF3273967DB3 wrong in cleaned too
      testresult < (median_testresult - 38.5))         ~ NA_real_, 
    testunits == "lb" &
      (testresult > (median_testresult + 99) | # choose 99 cm by default, # Problem when wrong first value FCC755E4F8722 (94.8) keep in case close to first dosinf
         testresult < (median_testresult - 99))      ~ NA_real_, # 91 for F00C16FEE3F74
    TRUE                                             ~ testresult
  )) %>% filter(str_detect(patientid, "FEF3273967DB3|F181950A26A19|F14EBD0D0FA81|F4A3B1B659865"))

weight1 <- weight %>% 
  # 3. Calculate median for each units and remove data outside range for each patients/unit
  # 3. help distinguishing lbs to kg for same patient when unit is NA
  group_by(patientid, testunits) %>%
  mutate(median_testresult = median(testresult, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(testresult_verified = case_when(       # Can improve by > 30 and > 10 from previous measure so need to create a gap var --------
    testunits == "kg" &
    (testresult > (median_testresult + 38.5) | # choose 38.5 cm by default could do 32 if not for F534F9C4FF968
       # 12 F2F31607FD6AD , FA7C8143AEE58 gain 37kg, F847A533FAE7E gain 38, FBE6E66AB2C95 gain 41, FC09E942CB0D7 gain 49
       # F5B18C17A28F5 and F98077E9119EC FA53D0F672C7D FCFF0EDFE278F, FEF3273967DB3 wrong in cleaned too
      testresult < (median_testresult - 38.5))         ~ NA_real_, 
    testunits == "lb" &
      (testresult > (median_testresult + 99) | # choose 99 cm by default, # Problem when wrong first value FCC755E4F8722 (94.8) keep in case close to first dosinf
         testresult < (median_testresult - 99))      ~ NA_real_, # 91 for F00C16FEE3F74
    TRUE                                             ~ testresult
  )) %>%  # FCC755E4F8722 keep the first value anyway, eliminate that F8CB4CAB1073A (has 96 lbs difference-median) ?
  # mutate(cal = testresult-median_testresult)
  # mutate(difference = (median_testresultcleaned)-testresultcleaned)

  # 4. Rescue testunits when testresult_verified is close 
  # Calculate max and min to limit value for each patient-Choose 20
  group_by(patientid, testunits) %>%
  mutate(max_kg = case_when(
    testunits == "kg" ~ max(testresult_verified, na.rm = TRUE) + 20 # F181950A26A19
  )) %>%
  mutate(min_kg = case_when(
    testunits == "kg" ~ min(testresult_verified, na.rm = TRUE) - 20
  )) %>%
  mutate(max_lbs = case_when(
    testunits == "lb" ~ max(testresult_verified, na.rm = TRUE) + 30 # F14EBD0D0FA81
  )) %>%
  mutate(min_lbs = case_when(
    testunits == "lb" ~ min(testresult_verified, na.rm = TRUE) - 30
  )) %>%
  ungroup() %>% 
  group_by(patientid) %>% 
  fill(max_kg, min_kg, max_lbs, min_lbs, .direction = "downup") %>% # min_kg, max_lbs
  ungroup() %>%

  mutate(testunits_rescue = case_when(
    is.na(testunits) &
      testresult_verified < max_kg        ~ "kg", # Good doesn't work with F16847515CAF4 , Perfect for F8A3DD629CDA6
    is.na(testunits) & # FE8DBCAC6DAFB
      testresult_verified > min_lbs      ~ "lb" # F004D683A1695, F10CAD0F4EC48, FC109A6F719E7, FD25ED170CF07, FE06FF7C3E108, F044601199C5D
  ))
weight1 %>% filter(str_detect(patientid, "FEF3273967DB3|F181950A26A19|F14EBD0D0FA81|F4A3B1B659865"))
```

Sometimes `testunits` was wrong or changed so try to rescue the unit within the same patient. For example these 2 patients have data in `kg` and earlier data with no units but doubled values so I assumed these data where on `lb.`

```{r}
weight2 <- weight1 %>% 
  group_by(patientid) %>% 
  # 5. Rescue testunits when testresult_verified is twice or /2 means that lb or kg 
  # if is 2 times the weight in kg add lb F14EBD0D0FA81
  mutate(testunits_rescue1 = case_when(
    is.na(testunits) &
      testresult_verified < (max_kg * 2.205) &
      testresult_verified > (min_kg * 2.205)         ~ "lb",
    is.na(testunits) &
      testresult_verified < (max_lbs / 2.205) &
      testresult_verified > (min_lbs / 2.205)        ~ "kg"
  )) %>% 
  ungroup() %>% 
  mutate(testunits = coalesce(testunits, testunits_rescue, testunits_rescue1)) %>% 
  select(c("patientid", "testdate", "testunits", "testresult", "testresult_verified", "testunitscleaned", "testresultcleaned", "weight1"))
weight1 %>% filter(str_detect(patientid, "F14EBD0D0FA81|F534F9C4FF968"))
```

The units are rescued here.
```{r}
weight2 %>% filter(str_detect(patientid, "F14EBD0D0FA81|F534F9C4FF968"))
```

Can finally recalculate weight.

<!-- F0027D3926C88 -->
```{r}
weight <- weight2 %>% 
  # 6.Fill up weight with new testresult_verified cleaned
  mutate(weight2 = case_when(
    is.na(testresultcleaned) &
      testunits == "kg"         ~ testresult_verified, # F0027D3926C88, FE8DBCAC6DAFB, F8A3DD629CDA6, F020C6A8B9E50, F16847515CAF4
    is.na(testresultcleaned) & # F044601199C5D
      testunits == "lb"         ~ (testresult_verified / 2.205), # FCB918617347C, F5E3D88009911, F14EBD0D0FA81, F16847515CAF4, FABC474931881
  )) %>% 
  mutate(weight = coalesce(weight1, weight2)) %>% 
  filter(!is.na(weight)) %>% 
  mutate(weight_units = "kg") %>% 
  select(c("patientid", weight_date = "testdate", "weight", "weight_units"))
weight
```

I can probably improve my cleaning the same way as height but beside that is it `r emo::ji("check")`
<br>

## AUC

I selected only AUC collected for the "taxel" and "platin" drugs but can have really high and low values.
```{r, rows.print=15, max.print=15}
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% arrange(relativeorderedamount) %>% select(patientid, relativeorderedamount, orderedamount, orderedunits, everything())
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% arrange(desc(relativeorderedamount)) %>% select(patientid, relativeorderedamount, orderedamount, orderedunits, everything())
```
I clean the data up by keeping the `relativeorderedamount`:  
- >= 50 & <= 175 for taxel  and with a unit in `mg/m2`  
- >= 2  & <= 6   for platin and with a unit in `AUC`
`r  emo::ji("question_mark")` But few example for taxel are remove for a `relativeorderedamount` at 36. Should we open it up a little more?
```{r}
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% 
  mutate(target_auc = case_when( # F7D7377578248 what for the patients who have order amount
    str_detect(drugname, "taxel") &
      relativeorderedunits == "mg/m2" &
      relativeorderedamount >= 50 &
      relativeorderedamount <= 175        ~ relativeorderedamount, # F21904E0D872F we should open more? F11B625154B62
    str_detect(drugname, "platin") &
      relativeorderedunits == "AUC" &
      relativeorderedamount >= 2 &
      relativeorderedamount <= 6          ~ relativeorderedamount, # F4E8C8717878B, FE3FF3A5AC465, F853DC24CD5E8
    TRUE                                  ~ NA_real_
  )) %>% filter(str_detect(patientid, "F21904E0D872F|F11B625154B62")) %>% select(patientid, drugname, relativeorderedamount,relativeorderedunits, target_auc, orderedamount, orderedunits, expectedstartdate)
```

`r  emo::ji("question_mark")` Same for platin?
```{r}
auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% 
  mutate(target_auc = case_when( # F7D7377578248 what for the patients who have order amount
    str_detect(drugname, "taxel") &
      relativeorderedunits == "mg/m2" &
      relativeorderedamount >= 50 &
      relativeorderedamount <= 175        ~ relativeorderedamount, # F21904E0D872F we should open more? F11B625154B62
    str_detect(drugname, "platin") &
      relativeorderedunits == "AUC" &
      relativeorderedamount >= 2 &
      relativeorderedamount <= 6          ~ relativeorderedamount, # F4E8C8717878B, FE3FF3A5AC465, F853DC24CD5E8
    TRUE                                  ~ NA_real_
  )) %>% filter(str_detect(patientid, "F4E8C8717878B|FE3FF3A5AC465|F853DC24CD5E8")) %>% select(patientid, drugname, relativeorderedamount,relativeorderedunits, target_auc, orderedamount, orderedunits, expectedstartdate)
```

`r  emo::ji("question_mark")` Also we will see later that a lot of patients don't have any `relativeorderedamount`. We talked and you advice to use the formula but I think that if we take the dose that a patient receive to reverse calculate the AUC we will automatically bias our results as we will create a AUC witch fit to the dose they received. Let me know if I misunderstood!!
`r  emo::ji("question_mark")` One other idea could be to calculate `relativeorderedamount` from `orderedamount` which look like it is just divided by bsa... Is that correct?
The resulted data
```{r AUC}
areaUC <- auc %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% 
  filter(!is.na(relativeorderedamount)) %>% 
  # select(c("patientid", "expectedstartdate", orderedamount, orderedunits, "relativeorderedamount",
  #          "relativeorderedunits", "drugname", iscanceled)) %>%
  mutate(target_auc = case_when( # F7D7377578248 what for the patients who have order amount
    # relativeorderedunits == "mg/kg" |
    #   relativeorderedunits == "mg" |
    #   relativeorderedunits == "m"         ~ NA_real_,
    str_detect(drugname, "taxel") &
      relativeorderedunits == "mg/m2" &
      relativeorderedamount >= 50 &
      relativeorderedamount <= 175        ~ relativeorderedamount, # F21904E0D872F we should open more? F11B625154B62
    str_detect(drugname, "platin") &
      relativeorderedunits == "AUC" &
      relativeorderedamount >= 2 &
      relativeorderedamount <= 6          ~ relativeorderedamount, # F4E8C8717878B, FE3FF3A5AC465, F853DC24CD5E8
    TRUE                                  ~ NA_real_
  )) %>%
  select(c("patientid", auc_date = "expectedstartdate", "target_auc", 
           auc_units = "relativeorderedunits", "drugname")) %>% 
  filter(!is.na(target_auc))
areaUC
```
Beside my question it is `r emo::ji("check")`
<br>

## Clinical
```{r}
clinical_data
```

I cleaned up `race` as I think we want `Hispanic` to be in the `White` category. I also calculate `month_at_os` with `diagnosisdate` and `followupdate`.
```{r}
clinical_data <- clinical_data %>% 
  mutate(race = case_when(
    str_detect(race, "Black")    ~ "Black",
    str_detect(race, "Hispanic") ~ "White",
    TRUE                         ~ race
  )) %>% 
  mutate(ethnicity = (str_remove(ethnicity, " or Latino"))) %>% 
  unite(raceeth1, c(race,ethnicity), sep = " ", remove = FALSE, na.rm = TRUE) %>% 
  mutate(vital = case_when(
    vitalstatus == 0     ~ "Alive",
    vitalstatus == 1     ~ "Dead"
  )) %>% 
  mutate(month_at_os = interval(start = diagnosisdate, end = followupdate)/
           duration(n=1, units = "months")) %>% 
  mutate(stagecat = na_if(stagecat, "Unk Stage")) %>% 
  mutate(across(.cols = c(histology, groupstage, tstage), ~na_if(., "Unknown/not documented")))
clinical_data
```
`r emo::ji("check")`
<br>

## Drugs

Here is the fun`r emo::ji("face_screaming_in_fear")`!  
The first cleanup step is to only keep the drugs which were **administered** to patients and removing the **maintenance** drugs, I also filtered **"taxel" and "platin"** drugs administered by **"intravenous" or "peritoneal"** route.  
`r  emo::ji("question_mark")` Done like that, I automatically remove the maintenance part within line which include non-maintenance drugs. 
<span style='color: red;'>I just want to make sure this is ok</span>

<!-- F71460509E771 is an example where it works perfectly     -->
Here are 2 examples where my code will remove the maintenance part within the first line but please take a look just to make sure.
```{r}
drugs %>% filter(str_detect(patientid, "F9E61D896E2FE|FA550FA15524A")) %>% filter(episodedatasource == "Administrations") %>% 
  arrange(patientid, episodedate) %>% select(patientid,  linenumber, ismaintenancetherapy, drugname,linename, episodedate,  everything())
```

Another thing I saw is that patients can have multiple records (rows) for the same drug (usually with different amount) on the same day.  
After our talk last time we decided to combined these multiple amount of drug given the same day. 
FYI, I found that on the Flatiron website:

*"There are a small number of patients with multiple records in the MedicationAdministration table for the same drug on the same day. We reviewed several instances where this occurs in the original EMR records and found, in each case, legitimate clinical reasons substantiating the multiple administrations per day. Although these instances were legitimate medication administrations, Flatiron Health data are generated from real-world clinical practice and are subject to miscoding and errors by the clinical oncology team, as are all EMR data. Below are three patient vignettes taken from our review of EMR notes, illustrating the variety of reasons why this may occur.  
For clinical reasons, a patient was given higher than normal dose which was listed as two administrations of the normal dose.  
Three administrations of the same drug were listed on a Friday. The EMR note stated that one dose was administered in the office and the patient was sent home with the two remaining doses for the weekend.   
Patient had an allergic reaction in the past and was administered multiple smaller doses in the same day as part of a desensitization protocol."*

It may be only 205 patients row like that (over 145517 total rows). Example here:
```{r}
drugs %>% 
  filter(episodedatasource == "Administrations" & ismaintenancetherapy == "FALSE" & !is.na(amount)) %>%
  filter(str_detect(drugname, "taxel|platin") & str_detect(route, "venous|peritoneal")) %>% 
  select(-c(ismaintenancetherapy, enhancedcohort, episodedatasource, drugcategory, detaileddrugcategory, route)) %>%
  filter(str_detect(patientid, "FCF86329BB40C|F4E6EE1910940")) %>% 
  select(patientid, episodedate, drugname, amount, everything())
```

In the `amount` variable I found `mg`, `ug`, `ml`, `Ea`, `AUC` and `NA`. They all look in the range (ug for ug and mg for everything else) but `AUC` has sometimes values < 10 (someone probably enter the real AUC!) so I removed these ones and convert `ug` amount into `mg`.
I also filtered out patients who had `issurgery` == "Yes" but no `surgerydate` associated (10 patients).
```{r drugs1}
drugs1 <- drugs %>% 
  # 1. Filtering----
  filter(episodedatasource == "Administrations" & ismaintenancetherapy == "FALSE" & !is.na(amount)) %>%
  filter(str_detect(drugname, "taxel|platin") & str_detect(route, "venous|peritoneal")) %>% 
  select(-c(ismaintenancetherapy, enhancedcohort, episodedatasource, drugcategory, detaileddrugcategory, route)) %>%
  group_by(patientid, linename, linenumber, linestartdate, lineenddate,episodedate, drugname, units) %>%
  summarise_at(vars(amount), paste, collapse = ";") %>% # don't do sum ti keep the multiple administration in df
  separate(col= amount, paste("amount_", 1:10, sep=""), sep = ";", extra = "warn",
           fill = "right") %>%
  purrr::keep(~!all(is.na(.))) %>%
  ungroup() %>% 
  mutate(amount_1 = as.numeric(amount_1), amount_2 = as.numeric(amount_2),
         amount_3 = as.numeric(amount_3), amount_4 = as.numeric(amount_4)) %>%
  mutate(amount = rowSums(select(.,amount_1:amount_4), na.rm = TRUE)) %>%
  # Found ug, ml, Ea, AUC and NA they are all good, but the AUC & < at 10 are wrong and ug need conversion
  mutate(amount =  case_when(
    units == "ug"               ~ amount / 10, # I checked all of them
    units == "AUC" &
      amount < 10               ~ NA_real_,
    TRUE                        ~ amount
  )) %>%

  # 3.bind with clinical----
  full_join(., clinical_data %>% select(-c(primaryphysicianid)),
            by = "patientid") %>% 
  # Limit to the patients we have date of surgery when had surgery
  filter(!(issurgery == "Yes" & is.na(surgerydate)))
```

Then the `linenumber` coding starts!!

`r  emo::ji("question_mark")` Just to make sure, when we talked, you said that we don't want to increase the line number when it is a change in cycle/drug (I don't think it happens often). 2 Examples here. The change of cycle type for the first patient happens in the second line and in the first line for the second patient. Is that still good?
```{r}
drugs1 %>% filter(str_detect(patientid, "F071D61E6FB93|F001443D8B85C")) %>% select(patientid, linename, linenumber, drugname, episodedate, everything())
```
`r  emo::ji("question_mark")` For this example, it is a change of drug and cycle type. I guess the patients didn't react well...but I still want to make sure it is ok.
```{r}
drugs1 %>% filter(str_detect(patientid, "F1775A21CC4DB")) %>% select(patientid, linename, linenumber, drugname, episodedate, everything())
```
<br>

>>>>

---
title: "Flatiron analysis"
author: "Christelle Colin-Leitzinger"
date: "3/17/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: flatly
    highlight: pygments
    df_print: paged
---

@@

# Analysis
***
<br>

@@

##
